---
title: "Week 14 IP Associatiative Rules"
author: "Njoki Mbugua"
date: "7/17/2021"
output: html_document
---
```{r}
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Defining the Question
## a) Specifying the Question
The objective of this project is to create association rules that will allow you to identify relationships between variables in the dataset.
## b) Defining the Metric for Success
Exhaustively performing associative analysis without any errors.

## c) Understanding the context
Working as a consultant Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).This project endeavors to  explore a recent marketing dataset by performing associative analysis to determine customer shopping patterns and later provide recommendations based on insights from the analysis.

## d) Recording the experimental design
Importing and reading the data
Data Cleaning
Perform associative analysis
Conclusions and recommendations

## e) Data Relevance
The data was provided by the company (http://bit.ly/SupermarketDatasetII). It has information on products purchased by customers over a span of time.

# Loading and previewing the data
```{r}
#First we call the arules library
#
library(arules)
library(tidyverse)
library(ggplot2)

```

```{r}
# We use the read.transactions function to load the data

url <- "http://bit.ly/SupermarketDatasetII"

df <- read.transactions ('http://bit.ly/SupermarketDatasetII',sep = ",")
df

```

```{r}
# Previewing the top 5 rows of the dataset
head(df)
```

```{r}
# Verifying the object's class
class(df)
```

```{r}
# Previewing our first 5 transactions

inspect(df[1:5])
```

```{r}
# Previewing the items that make up the dataset
items<-as.data.frame(itemLabels(df))
colnames(items) <- "Item"
head(items, 10)
```

```{r}
# Generating a summary of the dataset to view the most purchased items

summary(df)

```


```{r}
# Exploring the frequency of some articles 
itemFrequency(df[, 8:10],type = "absolute")

```
The top three frequently purchased items are black tea,blue berries and body spray.
```{r}
# Exploring the frequency of some articles
round(itemFrequency(df[, 8:10],type = "relative")*100,2)
```

```{r}
# Producing a chart of frequencies 
# Displaying top 10 most common items in the transactions dataset and the items whose relative importance is at least 10%
par(mfrow = c(1, 2))

# plot the frequency of items
itemFrequencyPlot(df, topN = 10,col="cyan")
itemFrequencyPlot(df, support = 0.1,col="blue")
```
The graphs present the top most items on the transactions.
```{r}
# Building a model based on association rules using the apriori function 
# We use Min Support as 0.001 and confidence as 0.8
rules <- apriori (df, parameter = list(supp = 0.001, conf = 0.8,maxlen=10))
rules

```
We obtain a set of 74 rules.

```{r}

summary(rules) 
```

```{r}
# Observing first 5 model rules 
inspect(rules[1:5])
```
Interpreting the top five rules:
a. If someone buys frozen smoothie and spinach, they are 88% likely to buy mineral water too
b. If someone buys bacon and pancakes, they are 81% likely to buy spaghetti too
c. If someone buys nonfat milk and turkey, they are 82% likely to buy mineral water too
d. If someone buys ground beef and nonfat milk, they are 86% likely to buy mineral water too
e. If someone buys frozen mushroom cream sauce and pasta, they are 95% likely to buy escalope too
```{r}
# Ordering these rules by a criteria such as the level of confidence
rules<-sort(rules, by="confidence", decreasing=TRUE)
inspect(rules[1:5])
```
Interpretation:
All the five rules have a confidence of 100.

```{r}
# creating a subset of rules for mineral water
milk.rules <- apriori(df, parameter = list(supp=0.001, conf=0.8),appearance = list(default="lhs",rhs="milk"))
```


```{r}
# Finding probability of how many customers buy milk
inspect(head(milk.rules))
```
Interpretation:
a. There is 82% chance that customers who bought black tea and smoothies also bought milk.
b. The probability of customers who bought meatballs and whole wheat pasta buying milk was 83%.
c. Customers who bought cake, meatballs and mineral water  will also by 100% chance purchase milk.

```{r}
# Visualizing the rules
# First filter rules with confidence greater than 0.5 or 50%
subrules <- rules[quality(rules)$confidence>0.5]
```

```{r}
library(arulesViz)
#Plot SubRules
plot(subrules)
```
From the plot we observe that rules with high lift have low support.

```{r}
subrules <- head(rules, n = 10, by = "confidence")

plot(subrules, method = "graph",  engine = "htmlwidget")
```
This graph shows associations between selected top ten items. Larger circles imply higher support, while red circles imply higher lift. 

