---
title: "Week 14 IP Feature Selection"
author: "Njoki Mbugua"
date: "7/17/2021"
output: html_document
---
```{r}
library(knitr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#  Defining the Question
## a) Specifying the Question
The objective of this project is to create perform feature selection by employing several to identify variables in the dataset that are relevant for prediction.

### b) Defining the Metric for Success
Successfully implement different feature selection methods.

### c) Understanding the context
Working as a consultant Data analyst at Carrefour Kenya and are currently undertaking a project that will inform the marketing department on the most relevant marketing strategies that will result in the highest no. of sales (total price including tax).This project endeavors to  explore a recent marketing dataset and perform feature selection to determine the features that are best suited for analysis.

### d) Recording the experimental design
Importing and reading the data
Data Cleaning
Implement feature selection techniques
Conclusions and recommendations

### e) Data Relevance
The data was provided by the company(http://bit.ly/CarreFourDataset).

```{r}
data = read.csv('http://bit.ly/CarreFourDataset',na.strings = "")
```


```{r}
head(data,n=5)
```

```{r}
# Installing the caret package and calling the library
suppressWarnings(
        suppressMessages(if
                         (!require(caret, quietly=TRUE))
                install.packages("caret")))
library(caret)
```


```{r}
# Calling the corrplot library
suppressWarnings(
        suppressMessages(if
                         (!require(corrplot, quietly=TRUE))
                install.packages("corrplot")))
library(corrplot)
```

```{r}
library(ggbiplot)
library("factoextra")
library(devtools)
library(plyr)
library(dplyr)
```

```{r}
# Selecting numeric column
df <- select_if(data, is.numeric)             
head(df)
```
```{r}
# Viewing the columns
colnames(df)
```
```{r}
# Removing columns with zero variance since we are getting an error that standard deviation is zero when calculating correlation
df2 <- df[,apply(df, 2, var, na.rm=TRUE) != 0]
```

```{r}
# Calculating the correlation matrix

corrMatrix <- cor(df2, method = "pearson",use = "complete.obs")
```


```{r}
# Find attributes that are highly correlated

highlyCorrelated <- findCorrelation(corrMatrix, cutoff=0.75)
```

```{r}
# Highly correlated attributes
# ---
# 
highlyCorrelated 
names(df2[,highlyCorrelated])
```
The cogs, total and tax columns are highly correlated.
```{r}
# Removing Redundant(highly correlated) Features 
# ---
# 
df3<-df2[-highlyCorrelated]
head(df3)
```

```{r}
# Performing our graphical comparison

par(mfrow = c(1, 2))
corrplot(corrMatrix, order = "hclust")
corrplot(cor(df3), order = "hclust")

```
The graphs show the correlation between variables.
Graph two shows the final features after dropping the highly correlated ones. Thus we can conclude that the most important features are gross income, quantity, unit price and rating.

`